Classifying galaxy spectra at $0.5<z<1$ with self-organizing maps

### 2018-03-25 sections0.08/answer_to_second_referee_comments.tex

We thank the reviewer for his/her helpful comments and apologize for the delay in returning a revised version of the manuscript. Large blocks of changed text in the manuscript appear in bold (we have also made minor grammatical and style changes which are not marked). The subsections of Section 4 have been reorganized both to address the referee's comments and make the flow more logical. Our responses to individual comments are in the text below, beginning with ---

@@

1. The authors did not address my comment about the significance of the work.  What was learned?  What method is advocated, and what is the evidence for its superiority?

The sentence that was added about 1D vs. 2D maps comes very late in the paper and is quite abstract, stating that the two methods are complementary.  The Introduction should establish the goals of the investigation so the reader understands the purpose of each experiment as well as its significance.

--- We have reorganized the last two paragraphs of Section 1 and added text to better describe the goal of the investigation: comparing different methods of classifying galaxy spectra against templates, with a focus on self-organizing maps.
We have added additional text to Sections 4.2 and 5 comparing 1D and 2D SOMs.
We have also added some text to Section 5 discussing the significance of the work.

The paper claims that "Compared to previous attempts to classify this sample, the self-organizing map approach was able to better classify the spectra which had similarities with more than one template." (abstract), but I cannot find a quantitative comparison with any previous results.  Figure 9 is presented as the result to compare with previous work, but since the previous results are not presented nor summarized, the reader can't tell how they differ.  The paper reports "more precision", but this is not defined; does it mean smaller error bars?  Stronger correlation?  How does this show "better classification"?  To support such a claim, this paper needs to select a performance metric and report on the values achieved in these experiments, compared to other methods.
--- We measure performance metrics in two areas: one set of metrics (an agreement score and the Fleiss kappa score) measures the degree to which the different classifications agree. The other set (silhouette score and Pearson correlation coefficient) measures the similarity between properties of galaxies assigned to the same class. The metrics are described in Section 3.4.

The results of this analysis are described in Sections 4.3 and 4.4. We have rewritten the abstract and summary to reflect the results.

I had suggested comparing the results to a simple nearest-neighbor classification baseline, but this was not addressed.
--- We apologize for missing this helpful suggestion in the previous version. This is now part of the analysis described in the paragraph above; the results appear in Section 4.3.

The comment about how SOMs provide an ordered clustering of the data (with respect to feature values), which k-means does not, is certainly true.  But the paper must then explain why this is important or valuable.
--- We have added text to the end of both Section 4.4 (formerly section 4.1) and Section 5 explaining the utility of an ordered clustering.

@@

On page 2, lines 29-31, the authors added this text: "An SOM can be trained without supervision and its results used to classify other data. Therefore, this method can be considered as (semi)-supervised." As written, the phrase "this method" seems to refer to SOMs (so the statement wouldn't be true - SOMs are fully unsupervised), but I think instead the authors' proposed method (of building an SOM and then applying it to classify new data, using labels on the SOM-grouped data) is what is meant.  Please clarify.
--- Changed to: 
    "The training of a self-organizing map is fully unsupervised.
     Using the resulting map as a template to classify other data requires labelling the output, which is why some groups consider self-organizing maps to be a (semi)-supervised method."

3. The inclusion of k-means helps give useful context to this work. Figure 10 shows a head-to-head comparison of k-means and SOM applied to the 12 templates using 4 clusters/nodes.  But I can't tell which one is better.  The text states that they are similar but "show discrepancies for spectra types between the extreme cases".  Which ones?  How can I tell from this figure?
--- The text describing the differences between the K-means and SOM classification of the templates (in particular, which templates are grouped together by the two methods) has been rewritten and moved to Section 4.1.2.

### 2018-03-25 main0.10.tex
Made new version to address latest (Feb 2018) ref report. Incorporated ref rept into tex files as comments.

\newcommand \boldit {\textbf{\textit{}}}
\newcommand \changed {\textbf{}}

@@

\import{sections0.09/}{intro0.09.tex}

\import{sections0.09/}{data0.09.tex}

\import{sections0.09/}{method0.09.tex}

\import{sections0.09/}{results0.09.tex}

%referee: Section 5 has been greatly improved and expanded; it now summarizes key observations that arose from the study and emphasizes its importance in further study of galaxy spectra.  In particular, the final paragraph (and the end of section 4.4) provides an excellent summary of why the SOM methods can be of unique value.  It should also include a statement about the results of other assessments (Fleiss kappa and silhouette) to provide a complete summary of the findings.

\section{Summary}
\label{sec: summary_SOMZ}

    Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.).
    In this work we presented a detailed comparison between different types of self-organizing maps and other methods used to classify galaxy spectra.
    Based on our experience we suggest some general guidelines for use of SOMs in galaxy spectral classification.
    If a broad and general classification is required, networks can have one dimension with a few neurons. 
    If one needs more detailed classifications, a higher number of neurons should be used.
    Since self-organizing maps do not include the uncertainty of input parameters, sometimes too much attention to detail can cause problems in classifications. 
    When using the SOM method, one should consider whether small differences between objects are physically meaningful when separating two groups from each other.

    We used SOMs to classify the template spectra of \citetalias{Kinney96}, made from galaxies with known morphological type, and created networks with different uses.
    By varying the size of the networks, we found the relative similarity between the \citetalias{Kinney96} template classes, which can be roughly ordered in one dimension by their amount of star formation.
     A one-dimensional network with 22 neurons was needed in order to
    separate all 12 \citetalias{Kinney96} spectra; we concluded that \citetalias{Kinney96} types B and E, and types SB1 and SB2, are very similar to each other.
    Two-dimensional networks allow more freedom in the galaxy spectral classification.
    Training $12\times 12$ neurons with the \citetalias{Kinney96} spectra, we found that the two dimensions of the resulting self-organized maps corresponded roughly to strength of star formation and amount of extinction. The one-dimensional self-organizing map ordering combined these two properties, with the highest-extinction template SB6 appearing between two less-extincted templates (Sb and Sc).

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
  This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although many of the spectra fall in between template classes.
     In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
 The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks: the supervised training method used by   \citetalias{Hossein12}    could not classify 37 out of 142 spectra in galaxy sample as matching one of the same set of templates.
    Comparing different classification methods, we found that spectra classified similarly by all methods were well-matched to quiescent templates (\citetalias{Kinney96} Sa and Sb). Spectra with emission lines or ultraviolet emission showed more variable classification results.
    Classification produced by self-organized maps showed closer correspondence than the supervised method results to the classification from a chi-squared match.
    Compared to classification using chi-squared minimization or the supervised neural network, the galaxies classed together by the self-organizing map had more similar spectra.

The relations between group-averaged properties of the sample galaxies using the SOM-based classification were found to be consistent with those from previous studies. 
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/star formation rate and specific star formation rate-stellar mass relations.
    Such an ordered classification does not naturally arise from other methods such as K-means clustering and supervised artificial neural networks.
    Ordered classification can be used to investigate underlying physical causes of changes in properties between groups and for this reason we conclude that self-organized maps can be a highly useful tool in exploratory analysis of astronomical spectra.

\section*{ACKNOWLEDGMENTS}
The authors thank the referee for insightful comments which helped us to improve the work.
The authors thank S. Lianou, A. Tammour, S.C. Gallagher, R.G. Abraham, M. Daley and A. Sigut for their useful comments. 
S.R. and P.B. also acknowledge research support from the Natural Sciences and Engineering Research Council of Canada.

@@

\import{sections0.09/}{apps_0.09.tex}
\end{document}

>>>>

% 22Jan2018: following 2 commands now work (they didn't before)
\newcommand{\boldit}[1]{\textbf{\textit{#1}}}
% for changed text to be bold: uncomment following line and comment out line after that
\newcommand{\changed}[1]{\textbf{#1}} 
% for changed text to be normal: comment out line above this, uncomment line below
%\newcommand{\changed}[1]{#1}

@@

\import{sections0.10/}{intro0.10.tex}

\import{sections0.10/}{data0.10.tex}

\import{sections0.10/}{method0.10.tex}

\import{sections0.10/}{results0.10.tex}

\import{sections0.10/}{summary0.10.tex}

\section*{ACKNOWLEDGMENTS}
The authors thank the referee for thorough and insightful comments which helped us to improve the work.
The authors thank S. Lianou, A. Tammour, S.C. Gallagher, R.G. Abraham, M. Daley and A. Sigut for their useful comments. 
S.R. and P.B. also acknowledge research support from the Natural Sciences and Engineering Research Council of Canada.

@@

\import{sections0.10/}{apps_0.10.tex}
\end{document}

### 2018-03-25 sections0.08/main0.08.tex
Made new version to address latest (Feb 2018) ref report. Incorporated ref rept into tex files as comments.

\newcommand \boldit {\textbf{\textit{}}}
\newcommand \changed {\textbf{}}

@@

\import{sections0.09/}{intro0.09.tex}

\import{sections0.09/}{data0.09.tex}

\import{sections0.09/}{method0.09.tex}

\import{sections0.09/}{results0.09.tex}

%referee: Section 5 has been greatly improved and expanded; it now summarizes key observations that arose from the study and emphasizes its importance in further study of galaxy spectra.  In particular, the final paragraph (and the end of section 4.4) provides an excellent summary of why the SOM methods can be of unique value.  It should also include a statement about the results of other assessments (Fleiss kappa and silhouette) to provide a complete summary of the findings.

\section{Summary}
\label{sec: summary_SOMZ}

    Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.).
    In this work we presented a detailed comparison between different types of self-organizing maps and other methods used to classify galaxy spectra.
    Based on our experience we suggest some general guidelines for use of SOMs in galaxy spectral classification.
    If a broad and general classification is required, networks can have one dimension with a few neurons. 
    If one needs more detailed classifications, a higher number of neurons should be used.
    Since self-organizing maps do not include the uncertainty of input parameters, sometimes too much attention to detail can cause problems in classifications. 
    When using the SOM method, one should consider whether small differences between objects are physically meaningful when separating two groups from each other.

    We used SOMs to classify the template spectra of \citetalias{Kinney96}, made from galaxies with known morphological type, and created networks with different uses.
    By varying the size of the networks, we found the relative similarity between the \citetalias{Kinney96} template classes, which can be roughly ordered in one dimension by their amount of star formation.
     A one-dimensional network with 22 neurons was needed in order to
    separate all 12 \citetalias{Kinney96} spectra; we concluded that \citetalias{Kinney96} types B and E, and types SB1 and SB2, are very similar to each other.
    Two-dimensional networks allow more freedom in the galaxy spectral classification.
    Training $12\times 12$ neurons with the \citetalias{Kinney96} spectra, we found that the two dimensions of the resulting self-organized maps corresponded roughly to strength of star formation and amount of extinction. The one-dimensional self-organizing map ordering combined these two properties, with the highest-extinction template SB6 appearing between two less-extincted templates (Sb and Sc).

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
  This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although many of the spectra fall in between template classes.
     In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
 The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks: the supervised training method used by   \citetalias{Hossein12}    could not classify 37 out of 142 spectra in galaxy sample as matching one of the same set of templates.
    Comparing different classification methods, we found that spectra classified similarly by all methods were well-matched to quiescent templates (\citetalias{Kinney96} Sa and Sb). Spectra with emission lines or ultraviolet emission showed more variable classification results.
    Classification produced by self-organized maps showed closer correspondence than the supervised method results to the classification from a chi-squared match.
    Compared to classification using chi-squared minimization or the supervised neural network, the galaxies classed together by the self-organizing map had more similar spectra.

The relations between group-averaged properties of the sample galaxies using the SOM-based classification were found to be consistent with those from previous studies. 
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/star formation rate and specific star formation rate-stellar mass relations.
    Such an ordered classification does not naturally arise from other methods such as K-means clustering and supervised artificial neural networks.
    Ordered classification can be used to investigate underlying physical causes of changes in properties between groups and for this reason we conclude that self-organized maps can be a highly useful tool in exploratory analysis of astronomical spectra.

@@

\import{sections0.09/}{apps_0.09.tex}
\end{document}

>>>>

\newcommand \boldit {\textbf{\textit{}}}

@@

\import{sections0.08/}{intro0.08.tex}

\import{sections0.08/}{data0.08.tex}

\import{sections0.08/}{method0.08.tex}

\import{sections0.08/}{results0.08.tex}

\section{Summary and Future Applications}
\label{sec: summary_SOMZ}

    Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.).
    {\bf In this work we presented a detailed comparison between different types of self-organizing maps and other methods used to classify galaxy spectra.
    Based on our experience we suggest some general guidelines for use of SOM in galaxy spectral classification.}
    If a broad and general classification is required, networks can have one dimension with a few neurons. 
    If one needs more detailed classifications, a higher number of neurons should be used.
    Since self-organizing maps do not include the uncertainty of input parameters, sometimes too much attention to detail can cause problems in classifications. 
    When using the SOM method, one should consider whether small differences between objects are physically meaningful when separating two groups from each other.

    We used SOMs to classify the template spectra of \citetalias{Kinney96}, made from galaxies with known morphological type, and created networks with different uses.
    By varying the size of the networks, we found the relative similarity between the \citetalias{Kinney96} template classes, which can be roughly ordered in one dimension by their amount of star formation.
     A one-dimensional network with 22 neurons was needed in order to
    separate all 12 \citetalias{Kinney96} spectra; we concluded that \citetalias{Kinney96} types B and E, and types SB1 and SB2, are very similar to each other.
    \textbf{Two-dimensional networks allow more freedom in the galaxy spectral classification.
    Training $12\times 12$ neurons with the \citetalias{Kinney96} spectra, we found that the two dimensions of the resulting self-organized maps corresponded roughly to strength of star formation and amount of extinction. The one-dimensional self-organizing map ordering combined these two properties, with the highest-extinction template SB6 appearing between two less-extincted templates (Sb and Sc).}

    \textbf{A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
  This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although many of the spectra fall in between template classes.
     In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
    The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks: the supervised training method used by   \citetalias{Hossein12} with the same templates and galaxy sample  could not classify 37 out of 142 spectra.
    Comparing different classification methods, we found that spectra classified similarly by all methods were well-matched to quiescent templates (\citetalias{Kinney96} Sa and Sb). Spectra with emission lines or ultraviolet emission showed more variable classification results.
    Classification produced by self-organized maps showed closer correspondence than the supervised method results to the classification from a chi-squared match.
    Compared to classification using chi-squared minimization or the supervised neural network, the galaxies classed together by the self-organizing map had more similar spectra.}

\textbf{The relations between group-averaged properties of the sample galaxies in the using the SOM-based classification were found to be consistent with those from previous studies. 
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/star formation rate and specific star formation rate-stellar mass relations.
    Such an ordered classification does not naturally arise from other methods such as K-means clustering and supervised artificial neural networks.
    Ordered classification can be used to investigate underlying physical causes of changes in properties between groups and for this reason we conclude that self-organized maps can be a highly useful tool in exploratory analysis of astronomical spectra.
}

@@

\import{sections0.08/}{apps_0.08.tex}
\end{document}

### 2018-03-25 sections0.08_for_editor/main0.08_for_editor.tex
Made new version to address latest (Feb 2018) ref report. Incorporated ref rept into tex files as comments.

\import{sections0.09/}{intro0.09.tex}

\import{sections0.09/}{data0.09.tex}

\import{sections0.09/}{method0.09.tex}

\import{sections0.09/}{results0.09.tex}

%referee: Section 5 has been greatly improved and expanded; it now summarizes key observations that arose from the study and emphasizes its importance in further study of galaxy spectra.  In particular, the final paragraph (and the end of section 4.4) provides an excellent summary of why the SOM methods can be of unique value.  It should also include a statement about the results of other assessments (Fleiss kappa and silhouette) to provide a complete summary of the findings.

\section{Summary}
\label{sec: summary_SOMZ}

    Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.).
    In this work we presented a detailed comparison between different types of self-organizing maps and other methods used to classify galaxy spectra.
    Based on our experience we suggest some general guidelines for use of SOMs in galaxy spectral classification.
    If a broad and general classification is required, networks can have one dimension with a few neurons. 
    If one needs more detailed classifications, a higher number of neurons should be used.
    Since self-organizing maps do not include the uncertainty of input parameters, sometimes too much attention to detail can cause problems in classifications. 
    When using the SOM method, one should consider whether small differences between objects are physically meaningful when separating two groups from each other.

@@

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
  This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although many of the spectra fall in between template classes.
     In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
 The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks: the supervised training method used by   \citetalias{Hossein12}    could not classify 37 out of 142 spectra in galaxy sample as matching one of the same set of templates.
    Comparing different classification methods, we found that spectra classified similarly by all methods were well-matched to quiescent templates (\citetalias{Kinney96} Sa and Sb). Spectra with emission lines or ultraviolet emission showed more variable classification results.
    Classification produced by self-organized maps showed closer correspondence than the supervised method results to the classification from a chi-squared match.
    Compared to classification using chi-squared minimization or the supervised neural network, the galaxies classed together by the self-organizing map had more similar spectra.

The relations between group-averaged properties of the sample galaxies using the SOM-based classification were found to be consistent with those from previous studies. 
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/star formation rate and specific star formation rate-stellar mass relations.
    Such an ordered classification does not naturally arise from other methods such as K-means clustering and supervised artificial neural networks.
    Ordered classification can be used to investigate underlying physical causes of changes in properties between groups and for this reason we conclude that self-organized maps can be a highly useful tool in exploratory analysis of astronomical spectra.

@@

\import{sections0.09/}{apps_0.09.tex}
\end{document}

>>>>

\import{sections0.08_for_editor/}{intro0.08.tex}

\import{sections0.08_for_editor/}{data0.08.tex}

\import{sections0.08_for_editor/}{method0.08.tex}

\import{sections0.08_for_editor/}{results0.08.tex}

\section{Summary and Future Applications}
\label{sec: summary_SOMZ}

    Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.).
    In this work we presented a detailed comparison between different types of self-organizing maps and other methods used to classify galaxy spectra.
    Based on our experience we suggest some general guidelines for use of SOM in galaxy spectral classification.
    If a broad and general classification is required, networks can have one dimension with a few neurons. 
    If one needs more detailed classifications, a higher number of neurons should be used.
    Since self-organizing maps do not include the uncertainty of input parameters, sometimes too much attention to detail can cause problems in classifications. 
    When using the SOM method, one should consider whether small differences between objects are physically meaningful when separating two groups from each other.

@@

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
  This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although many of the spectra fall in between template classes.
     In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
    The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks: the supervised training method used by   \citetalias{Hossein12} with the same templates and galaxy sample  could not classify 37 out of 142 spectra.
    Comparing different classification methods, we found that spectra classified similarly by all methods were well-matched to quiescent templates (\citetalias{Kinney96} Sa and Sb). Spectra with emission lines or ultraviolet emission showed more variable classification results.
    Classification produced by self-organized maps showed closer correspondence than the supervised method results to the classification from a chi-squared match.
    Compared to classification using chi-squared minimization or the supervised neural network, the galaxies classed together by the self-organizing map had more similar spectra.

The relations between group-averaged properties of the sample galaxies in the using the SOM-based classification were found to be consistent with those from previous studies. 
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/star formation rate and specific star formation rate-stellar mass relations.
    Such an ordered classification does not naturally arise from other methods such as K-means clustering and supervised artificial neural networks.
    Ordered classification can be used to investigate underlying physical causes of changes in properties between groups and for this reason we conclude that self-organized maps can be a highly useful tool in exploratory analysis of astronomical spectra.

@@

\import{sections0.08_for_editor/}{apps_0.08.tex}
\end{document}

### 2018-03-25 sections0.09/summary0.09.tex
Made new version to address latest (Feb 2018) ref report. Incorporated ref rept into tex files as comments.

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{amsmath}
\usepackage{hyperref}
\hypersetup{
    colorlinks=True,
    linkcolor=red,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=blue,
}
\usepackage{graphicx}
\usepackage{natbib}
\bibliographystyle{mn2e}
\usepackage{times}
\usepackage{float}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{color,soul}
\usepackage{import}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{amssymb}	% Extra maths symbols
\usepackage{multicol}        % Multi-column entries in tables
\usepackage{pgffor}
\usepackage{commath}	% Extra maths symbols

\newcommand \boldit {\textbf{\textit{}}}
\newcommand \changed {\textbf{}}

\newcommand \aaj {A\&A}
\newcommand \aarv {A\&ARv}%: Astronomy and Astrophysics Review (the)
\newcommand \aas{A\&AS}%: Astronomy and Astrophysics Supplement Series
\newcommand \afz {Afz}%: Astrofizika
\newcommand \aj {AJ}%: Astronomical Journal (the)
\newcommand \apss {Ap\&SS}%: Astrophysics and Space Science
\newcommand \apj {ApJ}
\newcommand \apjs {ApJS}%: Astrophysical Journal Supplement Series (the)
\newcommand \araa {ARA\&A} %: Annual Review of Astronomy and Astrophysics
\newcommand \asp {ASP Conf. Ser.}%: Astronomy Society of the Pacific Conference Series
\newcommand \azh {Azh}%: Astronomicheskij Zhurnal
\newcommand \baas {BAAS}%: Bulletin of the American Astronomical Society
\newcommand \mem {Mem. RAS}%: Memoirs of the Royal Astronomical Society
\newcommand \mnassa {MNASSA}%: Monthly Notes of the Astronomical Society of Southern Africa
\newcommand \mnras {MNRAS} %: Monthly Notices of the Royal Astronomical Society
%\newcommand {Nature}%(do not abbreviate)
\newcommand \pasj {PASJ}%: Publications of the Astronomical Society of Japan
\newcommand \pasp {PASP}%: Publications of the Astronomical Society of the Pacific
\newcommand \qjras {QJRAS}%: Quarterly Journal of the Royal Astronomical Society
\newcommand \mex {Rev. Mex. Astron. Astrofis.}%: Revista Mexicana de Astronomia y Astrofisica
%\newcommand {Science }%}%(do not abbreviate)
\newcommand \sva {SvA}%: Soviet Astronomy
\newcommand \aap {A\&A} %:Astronomy & Astrophysics
\newcommand \apjl {ApJL} %:The Astrophysical Journal Letters

\begin{document}
% TITLE
\defcitealias{Hossein12}{T12}
\defcitealias{Kinney96}{K96}
\defcitealias{Noll09}{N09}

\title[Classifying high-$z$ galaxy spectra]{Classifying galaxy spectra at $0.5<z<1$ with self-organizing maps}
%\author{rahmani.sahar}
\date{\today}
\author[S.~Rahmani, H.~Teimoorinia and P.~Barmby]{S.~Rahmani$^{1,2}$\thanks{E-mail:
srahma49@uwo.ca}, H.~Teimoorinia$^{3,4}$, P.~Barmby$^{1,2}$\\
$^{1}$Department of Physics $\&$ Astronomy, Western University, London, ON N6A 3K7, Canada\\
$^{2}$Centre for Planetary Science \& Exploration, Western University, London, ON N6A 3K7, Canada\\
$^{3}$Department of Physics $\&$ Astronomy, University of Victoria, Finnerty Road, Victoria, British Columbia, V8P 1A1, Canada\\
$^{4}$NRC Herzberg Astronomy and Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada }
\maketitle

\begin{abstract}
    The spectrum of a galaxy contains information about its physical properties.
    Classifying spectra using templates helps elucidate the nature of a galaxy's energy sources.
    In this paper, we investigate the use of self-organizing maps in classifying galaxy spectra against templates.
    We trained semi-supervised self-organizing map networks using a set of
    templates covering the wavelength range from far ultraviolet to near infrared. 
    The trained networks were used to classify the spectra of a sample of 142 galaxies with $0.5 < z < 1$ and the results compared to classifications performed using K-means clustering, a supervised neural network, and chi-squared minimization.
    Spectra corresponding to quiescent galaxies were more likely to be classified similarly by all methods while starburst spectra showed more variability.
    Compared to classification using chi-squared minimization or the supervised neural network, the galaxies classed together by the self-organizing map had more similar spectra.
     The class ordering provided by the one-dimensional self-organizing maps corresponds to an ordering in physical properties, a potentially important feature for the exploration of large datasets.
\end{abstract}
\begin{keywords} 
 galaxies: high-redshift, 
 galaxies: spectra, 
 methods: observational, 
 methods: statistical, 
 methods:data analysis
\end{keywords}

\import{sections0.09/}{intro0.09.tex}

\import{sections0.09/}{data0.09.tex}

\import{sections0.09/}{method0.09.tex}

\import{sections0.09/}{results0.09.tex}

%referee: Section 5 has been greatly improved and expanded; it now summarizes key observations that arose from the study and emphasizes its importance in further study of galaxy spectra.  In particular, the final paragraph (and the end of section 4.4) provides an excellent summary of why the SOM methods can be of unique value.  It should also include a statement about the results of other assessments (Fleiss kappa and silhouette) to provide a complete summary of the findings.

@@

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
  This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although many of the spectra fall in between template classes.
     In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
 The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks: the supervised training method used by   \citetalias{Hossein12}    could not classify 37 out of 142 spectra in galaxy sample as matching one of the same set of templates.
    Comparing different classification methods, we found that spectra classified similarly by all methods were well-matched to quiescent templates (\citetalias{Kinney96} Sa and Sb). Spectra with emission lines or ultraviolet emission showed more variable classification results.
    Classification produced by self-organized maps showed closer correspondence than the supervised method results to the classification from a chi-squared match.
    Compared to classification using chi-squared minimization or the supervised neural network, the galaxies classed together by the self-organizing map had more similar spectra.

@@

\section*{ACKNOWLEDGMENTS}
The authors thank the referee for insightful comments which helped us to improve the work.
The authors thank S. Lianou, A. Tammour, S.C. Gallagher, R.G. Abraham, M. Daley and A. Sigut for their useful comments. 
S.R. and P.B. also acknowledge research support from the Natural Sciences and Engineering Research Council of Canada.

\bibliographystyle{apalike}
\bibliography{ref_mining_h.bib}

\import{sections0.09/}{apps_0.09.tex}
\end{document}

>>>>

@@

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
  This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although many of the spectra fall in between template classes.
     In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
    The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks: the supervised training method used by \citetalias{Hossein12} could not classify 37 out of 142 spectra in galaxy sample as matching one of the same set of templates.

    \changed{
    Comparing different classification methods, we found that spectra classified similarly by all methods were well-matched to quiescent templates (\citetalias{Kinney96} Sa and Sb). 
    Spectra with emission lines or ultraviolet emission showed more variable classification results.
    Classification produced by self-organized maps showed closer correspondence than the supervised method results to the classification from a chi-squared match.
    Using the Fleiss kappa index to compare classifications showed that the K-means, supervised ANN and one-dimensional SOM are in fair agreement.
    One-dimensional self-organized-map based classification resulted in a 
    higher silhouette score for than classification using chi-squared minimization or the supervised neural network, indicating that the galaxies classed together by the SOM had more similar spectra.}

@@

### 2018-03-25 sections0.10/intro0.10.tex
Made new version to address latest (Feb 2018) ref report. Incorporated ref rept into tex files as comments.

@@

%Referee: Great list of questions near the end of section 1.  I recommend moving (or merging) the background information/explanation of the methods that are tested to section 3 so that this important content (the goals of the work) is featured more prominently and early on.
% DONE, but needs looked at again to make sure it makes sense.

@@

The goal of this work is to investigate the use of self-organizing maps in classifying galaxy spectra against templates, and compare them to other methods. 
The questions we want to answer are: 
(1) What is the degree of agreement between classifications produced by $\chi^2$ fitting, one- and two-dimensional self-organizing maps, K-means clustering and artificial neural networks? 
(2) Are certain spectral types more (or less) consistently classified by the different methods? 
(3) Which classification methods produce classes with the highest internal similarity ?
We analyze the same sample of 142 galaxies with $0.5 < z < 1$ as 
\citet[][hereafter \citetalias{Hossein12}]{Hossein12}, who classified the spectra using a supervised (multi-layer feed-forward) neural network method with the \citetalias{Kinney96} templates.
For each of the 142 galaxies, we find the best chi-squared fit among the 12 templates to use as a base for comparing the methods' performance.
We compute an agreement score for each method and also use the Fleiss Kappa index \citet{Fleiss71} to test interrater reliability 
between the methods.
To measure how well each classification method groups similar objects together, we
calculate the silhouette scores and dispersions in properties between galaxies classified in the same groups.

>>>>

% Referee comments:
%1a. The authors state (in their response, but not in the paper) that there is no known correct classification for these galaxies.  Please add this necessary context for the paper.  This positions the work in the world of of (semi-) supervision, since correct labels are not available.

%1b. The use of the chi-squared template assignment needs to be clearly explained and justified.  The author response states that this is a common method employed in the research community.  This too should be added to the paper to justify why it was chosen as a reference point (ideally with references to show its common usage by others).

%1c. If chi-squared minimization is the current most widely used method, then the paper needs to make a strong case for why to bother using a different method (SOMs) that only imperfectly approximates (i.e., does not have 100% agreement with) the chi-squared method.  I expect that the benefit is obtaining the insights provided from the analysis of the trained SOM structure, which cannot be achieved with the chi-squared method.  Make this explicit (up front, in the Introduction) so the reader understands why they might want to bother using this new method instead of chi-squared.

@@

@@

The goal of this work is to investigate the use of self-organizing maps in classifying galaxy spectra against templates, and compare them to other methods. 
The questions we want to answer are: 
(1) What is the degree of agreement between classifications produced by $\chi^2$ fitting, one- and two-dimensional self-organizing maps, K-means clustering and artificial neural networks? 
(2) Are certain spectral types more (or less) consistently classified by the different methods? 
(3) Which classification methods produce classes with the highest internal similarity ?
We analyze the same sample of 142 galaxies with $0.5 < z < 1$ as 
\citet[][hereafter \citetalias{Hossein12}]{Hossein12}, who classified the spectra using a supervised (multi-layer feed-forward) neural network method with the \citetalias{Kinney96} templates.
For each of the 142 galaxies, we find the best chi-squared fit among the 12 templates to use as a base for comparing the methods' performance.
We compute an agreement score for each method and also use the Fleiss Kappa index \citep{Fleiss71} to test interrater reliability 
between the methods.
To measure how well each classification method groups similar objects together, we
calculate the silhouette scores and dispersions in properties between galaxies classified in the same groups.
